Quantitative Benchmarking for Automatically Generated Conjectures

### 2017-10-27 TEBenchmarkPaper/article.tex
More citations, etc.

Automated theorem proving is an active area of research, with large problem sets
and regular competitions to prove as much as possible, as fast as possible.
These problem sets are an opportunity for theory exploration, as their
definitions and theorems can be used as a corpus, in the same way that Isabelle
libraries have been used in the past.

@@

\begin{itemize}
\item For each problem, there should be a clear distinction between the theorem
  to be proved and the definitions involved, such that the two can be easily and
  meaningfully separated. This rules out problem sets like SMTLIB, where many
  problems \emph{declare} a function, but leave its behaviour \emph{implicit} in
  the logical structure of the theorem statement.
\item Definitions should be \emph{strongly typed}, such that they can be
  translated into the native language of each ATE system (in this work, those
  languages are Haskell and Isabelle).
\item The problem set should be relevant to the desired domain. In our case, we
  desire higher-order functions and inductively defined types, which rules out
  first-order languages/logics.
\item Since it will act as our ground truth, the problem set should ideally
  contain every ``interesting'' conjecture involving its included
  definitions. Realistically, we should aim for each definition to appear in
  \emph{many} theorem statements; rather than each problem having unique
definitions.
\item The problem set should be as large as possible, for robustness of the
  resulting statistics.
\end{itemize}

@@

This neatly avoids the variance inherent in the summary statistics, and (with
enough samples) even approaches a normal distribution. Paired difference tests
are also robust to the censoring caused by timeouts: upper-bounding the run time
gives an upper bound to the (observable) differences, biasing our inference
towards the null hypothesis (indistinguishable performance) rather than
unfounded conclusions.

>>>>

Automated theorem proving is an active area of research, with large problem sets
and regular competitions to prove as much as possible, as fast as
possible\cite{pelletier2002development}. These problem sets are an opportunity
for theory exploration, as their definitions and theorems can be used as a
corpus, in the same way that Isabelle libraries have been used in the past.

@@

\begin{itemize}
\item For each problem, there should be a clear distinction between the theorem
  to be proved and the definitions involved, such that the two can be easily and
  meaningfully separated. This rules out problem sets like those of
  SMT-COMP~\cite{barrett2005smt}, where many problems involve uninterpreted
  functions, whose behaviour is \emph{implicit} in the logical structure of the
  theorem statement but not separately \emph{defined}.
\item Definitions should be \emph{strongly typed}, such that they can be
  translated into the native language of each ATE system (in this work, those
  languages are Haskell and Isabelle).
\item The problem set should be relevant to the desired domain. In our case, we
  desire higher-order functions and inductively defined types, which rules out
  first-order languages/logics (such as TPTP~\cite{sutcliffe2009tptp}).
\item Since it will act as our ground truth, the problem set should ideally
  contain every ``interesting'' conjecture involving its included
  definitions. Realistically, we should aim for each definition to appear in
  \emph{many} theorem statements; rather than each problem having unique
  definitions.
\item The problem set should be as large as possible, for robustness of the
  resulting statistics.
\end{itemize}

@@

This neatly avoids the variance inherent in the summary statistics, and (with
enough samples) approaches a normal distribution. Paired difference tests are
also robust to the censoring caused by timeouts: upper-bounding the run time
gives an upper bound to the (observable) differences, biasing our inference
towards the null hypothesis (indistinguishable performance) rather than
unfounded conclusions.
